IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design Under Nonlinear Harvester Model

### 2021-07-29 review/response_2.tex

		\begin{response}
			Thank you for sharing this paper. Instead of optimizing the phase shift of each IRS element, it tailored a tile-based two-stage optimization framework for large-scale IRS to enable a flexible performance-complexity tradeoff. The model also adapts well to practical IRS models accounting AoA and AoD. In contrast, our paper is more focused on waveform design and rate-energy tradeoff for one SWIPT user with practical co-located information decoder and energy harvester. We have updated the manuscript as follows.
			\begin{framed}
				In \cite{Xu2021}, the authors proposed a scalable resource allocation framework for SWIPT systems involving large-scale IRS, where the reflectors are grouped into tiles and the optimization process is divided into an offline mode-design stage and an online mode-selection stage to reduce overall complexity. It was concluded that the tile-based two-stage algorithm not only enables a flexible balance between performance and complexity but also adapts well to the physics-based IRS model accounting AoA and AoD. To the best of our knowledge, all existing IRS-aided SWIPT papers considered resource allocation and beamforming design for dedicated information and energy users in a single-carrier network. In this paper, we instead build our design based on a proper nonlinear harvester modeling that captures the dependency of the output DC power on both the power and shape of the input waveform, and marry the benefits of joint multi-carrier waveform and active beamforming optimization for SWIPT with the passive beamforming capability of IRS, to investigate the R-E tradeoff for one SWIPT user with practical co-located information decoder and energy harvester.
			\end{framed}
		\end{response}

		\begin{point}
			It could be better if the authors can provide some references for the parameter values chosen in the simulation part.
		\end{point}

		\begin{response}
			The system layout follows \cite{M-Wu2019} and we set the AP-UE distance to be larger than SWIPT literatures \cite{M-Clerckx2018b} and experiments \cite{M-Kim2020a} thanks to the IRS. For the center frequency, we consider \SI{2.4}{\GHz} due to its ubiquitous and license-free properties. Note that the path loss can be greatly reduced (and the R-E performance can be improved) if the system operates at much lower frequency, as considered in most IRS papers. In the previous manuscript, we made a mistake in the choice of reference path loss. As indicated by Comment~\ref{pt:3.3}, even the free-space path loss should be around \SI{40}{\dB} at \SI{1}{\meter}. To fix this issue, we updated the path loss model as IEEE TGn channel model D that is representative of small environments such as residential homes and small offices \cite{M-Erceg2004}. It has three clusters with \SI{50}{\ns} r.m.s. delay spread, and the path loss exponent is set to \num{2} (i.e., free-space model) up to breakpoint distance \SI{10}{\meter} and \num{3.5} onwards. This has a similar effect to most IRS literatures that assign a relatively large path loss exponent to the direct path and smaller ones to the incident and reflected paths to penalize the channel with large distance. As for the rectenna parameters, we forgot to mention that the parameters are taken from \cite{M-Clerckx2016a} and have added the reference to the revised manuscript.
		\end{response}

@@

		\begin{point}
			The figures are too small. For example, Fig.~\ref{M-fi:scaling_reflector}, the axis legend is not fully shown.
		\end{point}

		\begin{response}
			We apologize for the inconvenience. The figures have been amplified and the legends have been calibrated.
		\end{response}

		\begin{point}
			It is said that the reference path loss is \SI{-35}{\dB} at \SI{1}{\meter}, while for the center frequency at \SI{5.18}{\GHz}, even adopting the free space path loss model also will lead a much sever path loss at \SI{1}{\meter}. It is not clear how the authors calculate this.
			\label{pt:3.3}
		\end{point}

		\begin{response}
			We agree with the reviewer that the reference path loss used in the simulation is faulty and we sincerely apologize for this issue. The parameter \SI{-35}{\dB} was considered in some IRS literatures without specifying the operating frequency, and it was taken in the previous simulations without a second consideration. In the corrected manuscript, we instead consider a center frequency at \SI{2.4}{\GHz} that corresponds to a free-space path loss of \SI{-40}{\dB} at \SI{1}{\meter}. The path loss used in the simulation is taken from IEEE TGn channel model D \cite{M-Erceg2004} where the slope is set to \num{2} (i.e., free-space model) up to breakpoint distance \SI{10}{\meter} and \num{3.5} onwards. This has a similar effect to most IRS literatures that assign a relatively large path loss exponent to the direct path and smaller ones to the incident and reflected paths to penalize the channel with large distance.
		\end{response}

@@

		\begin{point}
			The authors considered the practical reflection coefficient beamforming in the simulation results. It is not clear how the authors obtain the discrete phase shift results. Direct quantization method and some other customized optimization techniques for discrete phase shifts may have significant performance gap, see \cite{Wu2020c}. Also, this work adopts the same uniform quantization as the above work, which thus needs to be clarified in the paper.
		\end{point}

		\begin{response}
			Thank you for the clarification. The discrete phase shift is obtained by quantizing the continuous phase shift obtained by Algorithm~\ref{M-al:bcd} to the uniform IRS codebook $\mathcal{C}_\phi = \{e^{j 2 \pi i / 2^b} \mid i = 1, \dots, 2^b\}$. We are aware that direct quantization can lead to performance loss compared with optimization over feasible phase shift set. In the revised manuscript, we have emphasized this point as follows.
			\begin{framed}
				On the other hand, since the practical reflection coefficient depends on the available element impedances, we consider a discrete uniform IRS codebook $\mathcal{C}_\phi = \{e^{j 2 \pi i / 2^b} \mid i = 1, \dots, 2^b\}$ and perform quantization on the continuous reflection coefficients returned by Algorithm~\ref{M-al:bcd} to reduce the circuit complexity and control overhead\footnotemark.
			\end{framed}
			\footnotetext{Note that this relax-then-quantize approach can bring notable performance loss compared with direct optimization over the discrete phase shift set, especially for a small $b$ (i.e., low-resolution IRS). The readers are referred to \cite{Wu2020c} for details.}
		\end{response}

		\begin{point}
			There are also some other practical phase shift models and nonlinear energy harvesting models such as saturation model for IRS-aided SWIPT in existing works, which may be discussed as related works such as amplitude-dependent phase shift model.
		\end{point}

		\begin{response}
			We have added reference \cite{Xu2021} in the updated manuscript that models the saturation effect of energy harvesters and physical properties of the IRS. Its contributions have been summarized in the manuscript as follows.
			\begin{framed}
				In \cite{Xu2021}, the authors proposed a scalable resource allocation framework for SWIPT systems involving large-scale IRS, where the reflectors are grouped into tiles and the optimization process is divided into an offline mode-design stage and an online mode-selection stage to reduce overall complexity. It was concluded that the tile-based two-stage algorithm not only enables a flexible balance between performance and complexity but also adapts well to the physics-based IRS model accounting AoA and AoD.
			\end{framed}
		\end{response}

@@

		\begin{point}
			Some early magazine papers and recent tutorial papers related to IRS-aided WPT/WIT may be further discussed.
		\end{point}

		\begin{response}
			We have incorporated the reference into the manuscript as follows.
			\begin{framed}
				\textellipsis The hardware architecture, design challenges and application opportunities of practical IRS are covered in \cite{Wu2020}.

				\textellipsis A recent tutorial paper \cite{Wu2021c} provided a comprehensive overview on IRS-aided wireless-powered networks, where channel estimation, resource allocation and practical constraints are discussed in detail.
			\end{framed}
		\end{response}
	\end{reviewer}

@@

	\begin{reviewer}
		\begin{point}
			What is the computational complexity of the proposed algorithm?
		\end{point}

		\begin{response}
			We have added complexity analysis in the revised manuscript. The reviewer is referred to Response~\ref{re:1.1} for details.
		\end{response}

		\begin{point}
			Some related works are worth citing, such as \cite{Yang2021}.
		\end{point}

		\begin{response}
			We have incorporated the reference into the manuscript as follows.
			\begin{framed}
				Deep reinforcement learning tools were also applied in \cite{Yang2021} to assist practical secure beamforming and reflection pattern design under QoS constraints for time-varying channels.
			\end{framed}
		\end{response}
	\end{reviewer}

### 2021-07-30 main.tex
:pencil: Update manuscript

			\emph{First}, we propose a novel IRS-aided SWIPT architecture based on joint waveform, active and passive beamforming design. To make an efficient use of the rectifier nonlinearity, we superpose a multi-carrier unmodulated power waveform (deterministic multisine) to a multi-carrier modulated information waveform and evaluate the performance under TS and PS receiving modes. The proposed joint waveform, active and passive beamforming architecture exploits the rectifier nonlinearity, a beamforming gain, and the channel selectivity across spatial and frequency domains to enlarge the achievable R-E region. This is the first paper to propose a joint waveform, active and passive beamforming architecture for IRS-aided SWIPT. By doing so, it is also the first paper to properly model the harvester nonlinearity (including its impact on both power and shape of the incoming waveform) and account for its crucial role in IRS-aided SWIPT.

@@

			\emph{Third}, we introduce two closed-form adaptive waveform schemes to avoid the exponential complexity of the GP algorithm. The Water-Filling (WF) strategy for modulated waveform and the SMF strategy for multisine waveform are combined in the time and power domains to facilitate practical SWIPT implementation. To accommodate the suboptimal waveform schemes, we modify the passive beamforming algorithm and characterize the R-E region by varying the duration ratio under TS mode and the combining and splitting ratios under PS mode. The proposed low-complexity designs achieve near-optimal R-E performance in different scenarios.

@@

		\begin{subsection}{Block Coordinate Descent}
			Based on the direct and cascaded CSIT, we iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:sca}, the active precoder $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b_n}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ and splitting ratio $\rho$ by Algorithm~\ref{al:gp}, until convergence. The steps are summarized in the BCD Algorithm~\ref{al:bcd}, which achieves local optimality with exponential computational complexity.

@@

			\begin{proposition}\label{pr:bcd}
				For any feasible initial point, the BCD Algorithm~\ref{al:bcd} is guaranteed to converge to local optimal points of the original waveform and beamforming problem~\eqref{op:original}.
			\end{proposition}

@@

		Fig.~\subref*{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \emph{First,} it is observed that increasing $N$ reduces the per-subband rate but boosts the harvested energy. It is because less power is allocated to each subband but more balanced DC terms are introduced to boost the harvested energy. On the other hand, Fig.~\subref*{fi:waveform_subband} sorts the modulated/multisine amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ for WPT in descending order. It demonstrates that a dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. This observation origins from the rectifier nonlinearity. Although both waveforms have equivalent second-order DC terms \eqref{eq:y_I2} and \eqref{eq:y_P2}, for the fourth-order terms \eqref{eq:y_I4} and \eqref{eq:y_P4}, the modulated waveform has $N^2$ monomials with a modulation gain of \num{2} while the multisine has $(2N^3+N)/3$ monomials. Hence, the benefit of multisine outstands for a sufficiently large $N$. \emph{Second,} the R-E region is convex for $N \in \{2,4\}$ and concave-convex for $N \in \{8,16\}$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e., a time sharing between the WPT point and the saddle SWIPT point obtained by PS (as the red curve in Fig.~\subref*{fi:re_subband}). Compared with the linear harvester model that requires no dedicated power waveform and always prefers PS, the rectifier nonlinearity enlarges the R-E region by favoring a different waveform and receiving mode, both heavily depend on the number of subbands. \emph{Third,} the LC-BCD algorithm achieves a very good R-E performance even if one-dimensional search is considered for $\delta=\rho$ over $[0,1]$. This conclusion is also verified in the following plots.

@@

	\begin{section}{Conclusion and Future Works}\label{se:conclusion_and_future_works}
		This paper investigated the R-E tradeoff of a single user employing practical receiving strategies in a novel IRS-aided multi-carrier MISO SWIPT system. Uniquely, we consider the joint waveform, active and passive beamforming design under rectifier nonlinearity to maximize the achievable R-E region. A three-stage local optimal BCD algorithm is proposed to solve the problem. In the first stage, the IRS phase shift is obtained by the SCA technique and eigen decomposition. In the second and third stages, the active precoder is derived in closed form, and the waveform amplitude and splitting ratio are optimized by the GP method. We also propose and combine closed-form adaptive waveform schemes with a modified passive beamforming strategy to formulate a low-complexity BCD algorithm that achieves near-optimal performance in different scenarios. Numerical results reveal significant R-E gains by modeling harvester nonlinearity in the passive beamforming and entire IRS-aided SWIPT design. Compared with active antennas, IRS elements cannot be designed independently across frequencies, but can integrate coherent combining and equal gain transmission in a fully passive manner to boost the array gain and power scaling order. However, a large number of reflecting elements may be required to compensate the double fading in the auxiliary link.

@@

		\begin{subsection}{Proof of Proposition~\ref{pr:bcd}}\label{ap:bcd}
			The objective function \eqref{ob:original} is non-decreasing over the iterations of Algorithm~\ref{al:bcd}, which is also upper-bounded due to the unit-modulus constraint \eqref{co:original_modulus} and the transmit power constraint \eqref{co:original_power}. Therefore, Algorithm~\ref{al:bcd} is guaranteed to converge. Since the IRS phase shift by Algorithm~\ref{al:sca} is local optimal, the active precoder by \eqref{eq:b_n} is global optimal, and the waveform amplitude and splitting ratio by Algorithm~\ref{al:gp} are local optimal, according to \cite{Grippo2000}, the limit point obtained by iteratively updating $\boldsymbol{\phi}$, $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$, $\boldsymbol{s}_{\mathrm{I/P}}$ and $\rho$ via Algorithm~\ref{al:bcd} is also local optimal to problem~\eqref{op:original}.
		\end{subsection}
	\end{appendix}

>>>>

			\emph{First}, we propose a novel IRS-aided SWIPT architecture based on joint waveform, active and passive beamforming design under the diode nonlinear model proposed in \cite{Clerckx2016a}. Although this tractable harvester model accurately reveals how the input power level and waveform shape influence the output DC power, it also introduces design challenges such as frequency compensation (i.e., components of different frequencies compensate and produce DC), waveform coupling (i.e., different waveforms jointly contribute to DC), and high-order objective function. To make an efficient use of the rectifier nonlinearity, we superpose a multi-carrier unmodulated power waveform (deterministic multisine) to a multi-carrier modulated information waveform and evaluate the performance under TS and PS receiving modes. The proposed joint waveform, active and passive beamforming architecture exploits the rectifier nonlinearity, a beamforming gain, and the channel selectivity across spatial and frequency domains to enlarge the achievable R-E region. This is the first paper to propose a joint waveform, active and passive beamforming architecture for IRS-aided SWIPT. By doing so, it is also the first paper to properly model the harvester nonlinearity (including its impact on both power and shape of the incoming waveform) and account for its crucial role in IRS-aided SWIPT.

@@

			\emph{Third}, we introduce two closed-form adaptive waveform schemes to avoid the exponential complexity of the GP algorithm. The Water-Filling (WF) strategy for modulated waveform and the SMF strategy for multisine waveform are combined in the time and power domains to facilitate practical SWIPT implementation. To accommodate the suboptimal waveform schemes, we modify the passive beamforming algorithm and characterize the R-E region by varying the duration ratio under TS mode and the combining and splitting ratios under PS mode. The proposed low-complexity designs achieve decent R-E performance in different scenarios.

@@

		\begin{subsection}{Block Coordinate Descent}
			Based on the direct and cascaded CSIT, we iteratively update the passive beamforming $\boldsymbol{\phi}$ by Algorithm~\ref{al:sca}, the active precoder $\boldsymbol{b}_{\mathrm{I/P},n}$, $\forall n$ by equation \eqref{eq:b_n}, and the waveform amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ and splitting ratio $\rho$ by Algorithm~\ref{al:gp}, until convergence. The steps are summarized in the BCD Algorithm~\ref{al:bcd}, whose computational complexity is exponential as inherited from Algorithm~\ref{al:gp}.

@@

			\begin{proposition}\label{pr:bcd}
				For any feasible initial point, the BCD Algorithm~\ref{al:bcd} is guaranteed to converge.
			\end{proposition}

@@

		Fig.~\subref*{fi:re_subband} illustrates the average R-E region versus the number of subband $N$. \emph{First,} it is observed that increasing $N$ reduces the per-subband rate but boosts the harvested energy. It is because less power is allocated to each subband but more balanced DC terms are introduced to boost the harvested energy. On the other hand, Fig.~\subref*{fi:waveform_subband} sorts the modulated/multisine amplitude $\boldsymbol{s}_{\mathrm{I/P}}$ for WPT in descending order. It demonstrates that a dedicated multisine waveform is unnecessary for a small $N$ but is required for a large $N$. This observation origins from the rectifier nonlinearity. Although both waveforms have equivalent second-order DC terms \eqref{eq:y_I2} and \eqref{eq:y_P2}, for the fourth-order terms \eqref{eq:y_I4} and \eqref{eq:y_P4}, the modulated waveform has $N^2$ monomials with a modulation gain of \num{2} while the multisine has $(2N^3+N)/3$ monomials. Hence, the benefit of multisine outstands for a sufficiently large $N$. \emph{Second,} the R-E region is convex for $N \in \{2,4\}$ and concave-convex for $N \in \{8,16\}$. This has the consequence that PS outperforms TS for a small $N$ and is outperformed for a large $N$. When $N$ is in between, the optimal strategy is a combination of both, i.e., a time sharing between the WPT point and the saddle SWIPT point obtained by PS (as the red curve in Fig.~\subref*{fi:re_subband}). For a relatively small $N$, the modulated waveform is used at both WIT and WPT point, and one can heuristically infer that no multisine waveform is needed for any R-E point (this is verified in simulation). In this case, the R-E region is convex, which aligns with the conventional linear harvester model (PS outperforms TS and dedicated power waveform is unnecessary). However, as $N$ becomes sufficiently large, multisine waveform further boosts the WPT point and creates some concavity in the high-power region, which accounts for the superiority of TS under nonlinear harvester model. In conclusion, the rectifier nonlinearity enlarges the R-E region by favoring a different waveform and receiving mode, both heavily depend on the number of subbands. \emph{Third,} the LC-BCD algorithm achieves a very good R-E performance even if one-dimensional search is considered for $\delta=\rho$ over $[0,1]$. This conclusion is also verified in the following plots.

@@

	\begin{section}{Conclusion and Future Works}\label{se:conclusion_and_future_works}
		This paper investigated the R-E tradeoff of a single user employing practical receiving strategies in a novel IRS-aided multi-carrier MISO SWIPT system. Uniquely, we consider the joint waveform, active and passive beamforming design under rectifier nonlinearity to maximize the achievable R-E region. A three-stage BCD algorithm is proposed to solve the problem. In the first stage, the IRS phase shift is obtained by the SCA technique and eigen decomposition. In the second and third stages, the active precoder is derived in closed form, and the waveform amplitude and splitting ratio are optimized by the GP method. We also propose and combine closed-form adaptive waveform schemes with a modified passive beamforming strategy to formulate a low-complexity BCD algorithm that achieves decent performance in different scenarios. Numerical results reveal significant R-E gains by modeling harvester nonlinearity in the passive beamforming and entire IRS-aided SWIPT design. Compared with active antennas, IRS elements cannot be designed independently across frequencies, but can integrate coherent combining and equal gain transmission in a fully passive manner to boost the array gain and power scaling order. However, a large number of reflecting elements may be required to compensate the double fading in the auxiliary link.

@@

		\begin{subsection}{Proof of Proposition~\ref{pr:bcd}}\label{ap:bcd}
			The objective function \eqref{ob:original} is non-decreasing over the iterations, because the IRS phase shift by Algorithm~\ref{al:sca} is local optimal, the active precoder by \eqref{eq:b_n} is global optimal, and the waveform amplitude and splitting ratio by Algorithm~\ref{al:gp} are local optimal. \eqref{ob:original} is also upper-bounded due to the unit-modulus constraint \eqref{co:original_modulus} and the transmit power constraint \eqref{co:original_power}. Therefore, Algorithm~\ref{al:bcd} is guaranteed to converge. However, it may not converge to optimal points since variables are coupled in constraint~\eqref{co:original_rate} \cite{Grippo2000}.
		\end{subsection}
	\end{appendix}
